---
title: "Racial Health Disparities in the USA"
output: html_document
---
**Intro**
<p>The domain of the project focuses on disparities in demographics in relation to healthcare in America. By analyzing data sets on health care resources and racial demographics by state we can observe the correlation between minority groups and their access to healthcare. This project proposal aims to seek resources that most affect racial groups to find better solutions to healthcare across the US. </p>
<p>**Summary**</p>
<p>With the functions, we were able to find out that 
  
<p>Some observations we made were that states that are larger tended to have more hospitals/medical resources available to them as the populous states have higher number of citizens seeking medical attention.

<p>**Table**</p>
```{r echo = FALSE, message = FALSE, warning = FALSE}
hc <- read.csv("https://www.ahrq.gov/sites/default/files/wysiwyg/chsp/compendium/chsp-compendium-2018.csv", stringsAsFactors = FALSE)

summary <- hc %>% 
  group_by(health_sys_state) %>% 
  summarize(beds = sum(sys_beds, na.rm = T),
            medics = sum(total_mds, na.rm = T),
            hospitals = sum(hosp_cnt, na.rm = T),
            medical_groups = sum(grp_cnt, na.rm = T),
            discharge = sum(sys_dsch, na.rm = T))
```

The aggregate table summarizes the different 

<p>**Pie Chart**</p>
```

```

<p>**Map**: Number of MD in each state</p>
```{r echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(usmap)
library(ggplot2)

#Import Tables
hc <- read.csv("https://www.ahrq.gov/sites/default/files/wysiwyg/chsp/compendium/chsp-compendium-2018.csv", stringsAsFactors = FALSE)

# sum of medics by state
sum_total_mds <- hc %>%
  group_by(health_sys_state) %>%
  summarize(sum_total_mds = sum(total_mds)) %>% 
  mutate(fips = fips(health_sys_state))

plot_usmap(regions = "states", data = sum_total_mds, values = "sum_total_mds", color = "blue") + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "number of mds", label = scales::comma
  ) + theme(legend.position = "right")
```

This chart is a density map of the U.S. states shaded by the number of physicians in health systems in each state. States that are shaded with a lighter color have fewer physicians and states that are shaded with a darker blue have more physicians. The chart was intended to visualize how access to healthcare resources (e.g. physicians) varies in different states and across different regions. From this chart, we see that California has the most number of physicians, followed by Northeastern states, such as New York and Pennsylvania. On the other hand, the mountain states, including Wyoming, Nevada, and Idaho, do not have many physicians.

<p>**Scatter Plot**</p>
```{r echo = FALSE, message = FALSE, warning = FALSE}
# Import Tables
hc <- read.csv("https://www.ahrq.gov/sites/default/files/wysiwyg/chsp/compendium/chsp-compendium-2018.csv", stringsAsFactors = FALSE)

# sum of beds by state
sum_total_beds <- hc %>%
  group_by(health_sys_state) %>%
  summarize(sum_total_beds = sum(sys_beds)) %>% 
  rename(state = health_sys_state)

# sum of discharges by state
sum_total_dsch <- hc %>%
  group_by(health_sys_state) %>%
  summarize(sum_total_dsch = sum(sys_dsch)) %>% 
  rename(state = health_sys_state)

# join data frames
join_beds_dsch <- left_join(sum_total_beds, sum_total_dsch)

# scatterplot of proportion of minorities to health care resources
ggplot(data= join_beds_dsch) +
  geom_point(mapping = aes(
    x = sum_total_dsch , y = sum_total_beds, col = state
  )) 
```
<p>This scatter plot indicates the correlation between the total number of hospital beds to the total number of discharges by state. This evaluates patterns and the number of beds in each state and shows that higher level of bed capacity directly relates to higher number of patient discharges. Grouping each category by state also shows what each state's average bed capacity to discharge rates as well.</p>
